---
title: "MFStocks Analysis"
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = FALSE, warning = FALSE, tidy = TRUE)
options(width=120)
```

# Read Data

```{r Data, include=FALSE}
setwd("~/Projects/mfstocks/code_and_data/analysis") # data path
stocks60 <- read.table("data_60.csv", header=TRUE, stringsAsFactors=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)  # READ DATA
```


```{r Data, include=FALSE}
stocks60$dt_60 <- as.POSIXct(stocks60$dt_60, tz="EST") # CONVERT DT TO POSIX
col60 <- c("season_workday_60", "season_month_60", "tf2_60") 
stocks60[col60] <- lapply(stocks60[col60], as.factor) # CONVERT CAT VARs TO FACTORs
```

# Preprocessing

Load packages
```{r Load packages}
library(RcmdrMisc)
library(dplyr) # %>% function
library(reshape2) # melt() 
library(lubridate) # date()
```

Create a morality index, using the sum of the sentiment-weighted probabilities of each of the foundations. 
```{r Create morality index}
stocks60$morality = 
  stocks60$care_p_60*stocks60$care_sent_60 + 
  stocks60$fairness_p_60*stocks60$fairness_sent_60 +
  stocks60$loyalty_p_60*stocks60$loyalty_sent_60 +
  stocks60$authority_p_60*stocks60$authority_sent_60 +
  stocks60$sanctity_p_60*stocks60$sanctity_sent_60
```


Take the difference of the stock closing values to see the changes in stock market movement. The variance is quite unstable so we then log the difference to try to stabilize it.  
```{r Make stationary}
# Stocks 
stocks60$stocks_60_diff = stocks60$stocks_60 %>% diff() %>% append(NA, 0) # Difference for moving averages
stocks60$stocks_60_diff_ln = stocks60$stocks_60 %>% log() %>% diff() %>% append(NA, 0) # Log then difference for variance stabilization
stocks60$day_count = stocks60$dt_60 %>% date() %>% as.integer() - stocks60$dt_60 %>% date() %>% as.integer() %>% min() + 1 # produce day_count predictor

# Morality predictors 
foundations <- c("care","fairness","loyalty","authority","sanctity")
stocks60[foundations] <- stocks60[6:10]*stocks60[11:15]

stocks60$morality_diff =  stocks60$morality %>% diff() %>% append(NA, 0) 
stocks60$care_diff = stocks60$care %>% diff() %>% append(NA, 0)
stocks60$fairness_diff = stocks60$fairness %>% diff() %>% append(NA, 0) 
stocks60$loyalty_diff = stocks60$loyalty %>% diff() %>% append(NA, 0)
stocks60$authority_diff = stocks60$authority %>% diff() %>% append(NA, 0) 
stocks60$sanctity_diff = stocks60$sanctity %>% diff() %>% append(NA, 0)
```

```{r}
stocks60 %>% colnames()

```

Next, add lags and leads to all the predictors, including the morality index (1), all moral probabilities for each foundation (5), all moral sentiments for each foundation (5), and all sentiment-weighted probabilities for each foundation (5). We started with 25 variables in the stocks data frame. We should end up with 57 variables. 
```{r Lag and lead predictors}
# morality
stocks60$morality_lag = lag(stocks60$morality_diff)
stocks60$morality_lead = lead(stocks60$morality_diff)

stocks60[paste(foundations, "_lag", sep="")] = lag(stocks60[27:31])
stocks60[paste(foundations, "_lead", sep="")] = lead(stocks60[27:31])
```

Now that we have all of our predictor variables ready, we can save the processed data. 
```{r}
dir.create("processed_data")
write.csv(stocks60, "processed_data/stocks60.csv")
```

Make time series and plot. We can see that the log minimized the variance somewhat, but it is still somewhat non-stationary. Perhaps future analyses can use a GARCH model to further process the data, but for the sake of our analysis, it is ok. 
```{r Make time series}
stocks60ts = ts(stocks60) # MAKE TIME SERIES
plot(stocks60ts[,"stocks_60"]) # PLOT INITIAL DATA
plot(stocks60ts[,"stocks_60_diff"]) # PLOT DIFFERENCED DATA
plot(stocks60ts[,"stocks_60_diff_ln"]) # PLOT LOG DATA

plot(stocks60ts[,"morality_diff"])
plot(stocks60ts[,"care_diff"])
plot(stocks60ts[,"fairness_diff"])
plot(stocks60ts[,"loyalty_diff"])
plot(stocks60ts[,"authority_diff"])
plot(stocks60ts[,"sanctity_diff"])
```



Plot ACF and PACF
```{r Serial Dependencies; ACF & PACF}
plot(stocks60ts[,"stocks_60"]) # PLOT INITIAL DATA

acf(stocks60$stocks_60, lag.max = NULL, type = c("correlation"), plot = TRUE, na.action = na.pass) # ACF FOR NON-TRANSFORMED DATA
acf(stocks60$stocks_60, lag.max = NULL, type = c("partial"), plot = TRUE, na.action = na.pass) # PACF FOR NON-TRANSFORMED DATA

acf(stocks60$stocks_60_diff_ln, lag.max = NULL, type = c("correlation"), plot = TRUE, na.action = na.pass) # ACF FOR TRANSFORMED DATA
acf(stocks60$stocks_60_diff_ln, lag.max = NULL, type = c("partial"), plot = TRUE, na.action = na.pass) # PACF FOR TRANSFORMED DATA

plot(stocks60ts[,"stocks_60_diff_ln"]) # 
```

```{r Serial Dependencies; ACF & PACF}
plot(stocks60ts[,"morality"]) # PLOT INITIAL DATA

acf(stocks60$morality, lag.max = NULL, type = c("correlation"), plot = TRUE, na.action = na.pass) # ACF FOR NON-TRANSFORMED DATA
acf(stocks60$morality, lag.max = NULL, type = c("partial"), plot = TRUE, na.action = na.pass) # PACF FOR NON-TRANSFORMED DATA

acf(stocks60$morality_diff, lag.max = NULL, type = c("correlation"), plot = TRUE, na.action = na.pass) # ACF FOR TRANSFORMED DATA
acf(stocks60$morality_diff, lag.max = NULL, type = c("partial"), plot = TRUE, na.action = na.pass) # PACF FOR TRANSFORMED DATA

plot(stocks60ts[,"morality_diff"]) # 
```

# Get Decriptives

```{r Descriptive statistics}
library(pastecs)
cols <- c("morality","care","fairness","loyalty","authority","sanctity", "stocks_60", "stocks_60_diff","stocks_60_diff_ln")
stocks60[,cols] %>% 
  stat.desc()
```


# Removing Outliers

Values above Q3 + 1.5xIQR or below Q1 - 1.5xIQR are considered as outliers. Values above Q3 + 3xIQR or below Q1 - 3xIQR are considered as extreme points (or extreme outliers). 
```{r Outliers}
library(rstatix)
outliers <- stocks60 %>% 
  identify_outliers(c(morality))
print(paste("Number of outliers: ", nrow(outliers)))

stocks60_outrm <-  stocks60 %>% dplyr::filter(!(morality %in% outliers$morality))
print(paste("Before outlier removal: ", nrow(stocks60))) 
print(paste("After outlier removal: ", nrow(stocks60_outrm))) 
print(paste("Difference: ", nrow(stocks60) - nrow(stocks60_outrm)))
```


# Cross correlation

The lag k value returned by ccf(x, y) estimates the correlation between x[t+k] and y[t]. 
```{r}
plot_TF = TRUE

lag_max = 5
ylab = "ccf"

png("figs/ccf.png", width = 5, height=4, units='in', res=300)


par(mfrow=c(2,3))
ccf(stocks60_outrm$morality_diff, stocks60_outrm$stocks_60_diff_ln, na.action = na.exclude, lag.max = lag_max, plot= plot_TF, ylab=ylab, main="Stocks & Morality Lags")
ccf(stocks60_outrm$care_diff, stocks60_outrm$stocks_60_diff_ln, na.action = na.exclude, lag.max = lag_max, plot= plot_TF, ylab=ylab, main="Stocks & Care Lags")
ccf(stocks60_outrm$fairness_diff, stocks60_outrm$stocks_60_diff_ln, na.action = na.exclude, lag.max = lag_max, plot= plot_TF, ylab=ylab, main="Stocks & Fairness Lags")
ccf(stocks60_outrm$loyalty_diff, stocks60_outrm$stocks_60_diff_ln, na.action = na.exclude, lag.max = lag_max, plot= plot_TF, ylab=ylab, main="Stocks & Loyalty Lags")
ccf(stocks60_outrm$authority_diff, stocks60_outrm$stocks_60_diff_ln, na.action = na.exclude, lag.max = lag_max, plot= plot_TF, ylab=ylab, main="Stocks & Authority Lags")
ccf(stocks60_outrm$sanctity_diff, stocks60_outrm$stocks_60_diff_ln, na.action = na.exclude, lag.max = lag_max, plot= plot_TF, ylab=ylab, main="Stocks & Sanctity Lags")
dev.off()
```

# Fitting and Evaluating Models
```{r}
library(jtools) # summ()
library(lme4) # lme models
library(interactions)
```

Generate models

```{r}
stocks60_outrm <- stocks60_outrm %>% na.omit()
```

Of the following cells (lagged predictors, or leading predictors), run ONLY 1.

Lagged predictors
```{r Lagged predictors}
# Lag
## Linear regression
model.lm <- 
  lm(stocks_60_diff_ln ~ tf2_60 * morality_lag,
     data = stocks60_outrm)

model.lm.foundations <- 
  lm(stocks_60_diff_ln ~ tf2_60 * (care_lag +fairness_lag +loyalty_lag +authority_lag +sanctity_lag), 
     data = stocks60_outrm)

## Random intercept with fixed mean (null)
model.lme.null <-
  lmer(stocks_60_diff_ln ~ 1 + (1|day_count), 
       data = stocks60_outrm, REML=TRUE)

# Correlated random slopes
model.lme <- 
  lmer(stocks_60_diff_ln ~ 
         (1 + season_intraday_60 + tf2_60*morality_lag + (1 + season_intraday_60 | day_count)), 
       data = stocks60_outrm, REML = TRUE)

model.lme.foundations <- 
  lmer(stocks_60_diff_ln ~
         1 + season_intraday_60 + tf2_60*(care_lag + fairness_lag + loyalty_lag + authority_lag + sanctity_lag) + ( 1 + season_intraday_60 | day_count), 
       data = stocks60_outrm, REML = TRUE)
```

Leading predictors
```{r Leading predictors}
# # Lead
# 
# ## Linear regression
# model.lm <-
#   lm(stocks_60_diff_ln ~ tf2_60 * morality_lead,
#      data = stocks60_outrm)
# 
# model.lm.foundations <- 
#   lm(stocks_60_diff_ln ~ tf2_60 * (care_lead +fairness_lead +loyalty_lead +authority_lead +sanctity_lead), 
#      data = stocks60_outrm)
# 
# ## Random intercept with fixed mean (null)
# model.lme.null <-
#   lmer(stocks_60_diff_ln ~ 1 + (1|day_count), 
#        data = stocks60_outrm, REML=TRUE)
# 
# # Correlated random slopes
# model.lme <- 
#   lmer(stocks_60_diff_ln ~ 
#          (1 + season_intraday_60 + tf2_60*morality_lead + (1 + season_intraday_60 | day_count)), 
#        data = stocks60_outrm, REML = TRUE)
# 
# model.lme.foundations <- 
#   lmer(stocks_60_diff_ln ~  
#          (1 + season_intraday_60 + tf2_60*(care_lead + fairness_lead + loyalty_lead + authority_lead + sanctity_lead) + ( 1 + season_intraday_60 | day_count)), 
#        data = stocks60_outrm, REML = TRUE)
```



```{r Model Comparisons: morality and foundations}
anova(model.lm, model.lm.foundations)
anova(model.lme.null, model.lme, model.lme.foundations)
```

```{r Model Comparisons: linear and mixed-effects}
anova(model.lme, model.lm, type="Chisq")
anova(model.lme.foundations, model.lm.foundations, type="Chisq")
```

# ANOVA results

```{r ANOVA}
Anova(model.lm, type="III", test="F")
Anova(model.lm.foundations, type="III", test="F")
Anova(model.lme.null, type="III", test="F")
Anova(model.lme, type="III", test="F")
Anova(model.lme.foundations, type="III", test="F")
```


```{r Standardized Coefficients}
summ(model.lme, scale=TRUE, transform.response=TRUE, confint=TRUE, digits=2)
summ(model.lme.foundations, scale=TRUE, transform.response=TRUE, confint=TRUE, digits=2)
```

# Plot interaction variables

```{r}
# define models
lag_mor <- lmer(stocks_60_diff_ln ~ (1 + season_intraday_60 + tf2_60*morality_lag + (1 + season_intraday_60 | day_count)), 
       data = stocks60_outrm, REML = TRUE)
lag_f <- lmer(stocks_60_diff_ln ~1 + season_intraday_60 + tf2_60*(care_lag + fairness_lag + loyalty_lag + authority_lag + sanctity_lag) + ( 1 + season_intraday_60 | day_count), 
       data = stocks60_outrm, REML = TRUE)
lead_mor <- lmer(stocks_60_diff_ln ~ (1 + season_intraday_60 + tf2_60*morality_lead + (1 + season_intraday_60 | day_count)), 
       data = stocks60_outrm, REML = TRUE)
lead_f <- lmer(stocks_60_diff_ln ~  (1 + season_intraday_60 + tf2_60*(care_lead + fairness_lead + loyalty_lead + authority_lead + sanctity_lead) + ( 1 + season_intraday_60 | day_count)), 
       data = stocks60_outrm, REML = TRUE)

# define plots
lag_m <- interact_plot(model = lag_mor, pred = morality_lag, modx = tf2_60, interval = TRUE, x.label="Morality", y.label="Market Movement", modx.labels = c("High","Low"), legend.main="Economic \n Period") + ylim(-0.02,0.02) + xlim(-0.02,0.02)  +rremove("ylab")+ theme(axis.text.x = element_text(size=8))
lag_f1 <- interact_plot(model = lag_f, pred = authority_lag, modx = tf2_60, interval = TRUE, x.label="Authority", y.label="Market Movement", modx.labels = c("High","Low"), legend.main="Economic Period") + ylim(-0.02,0.02)+ xlim(-0.005,0.003) + rremove("legend")+rremove("ylab")+ theme(axis.text.x=element_text(size=8), axis.text.y=element_blank())
lag_f2 <- interact_plot(model = lag_f, pred = sanctity_lag, modx = tf2_60,interval = TRUE, x.label="Sanctity", y.label="Market Movement", modx.labels = c("High","Low"), legend.main="Economic Period") + ylim(-0.02,0.02)+ xlim(-0.005,0.003)+ rremove("legend")+rremove("ylab")+ theme(axis.text.x=element_text(size=8),axis.text.y=element_blank())

lead_m <- interact_plot(model = lead_mor, pred = morality_lead, modx = tf2_60, interval = TRUE, x.label="Morality", y.label="Market Movement", modx.labels = c("High","Low"), legend.main="Economic \n Period") + ylim(-0.02,0.02) + xlim(-0.02,0.02)  +rremove("ylab")+ theme(axis.text.x = element_text(size=8))
lead_f1 <- interact_plot(model = lead_f, pred = authority_lead, modx = tf2_60, interval = TRUE, x.label="Authority", y.label="Market Movement", modx.labels = c("High","Low"), legend.main="Economic Period") + ylim(-0.02,0.02)+ xlim(-0.005,0.003) + rremove("legend")+rremove("ylab")+ theme(axis.text.x=element_text(size=8), axis.text.y=element_blank())
lead_f2 <- interact_plot(model = lead_f, pred = sanctity_lead, modx = tf2_60,interval = TRUE, x.label="Sanctity", y.label="Market Movement", modx.labels = c("High","Low"), legend.main="Economic Period") + ylim(-0.02,0.02)+ xlim(-0.005,0.003)+ rremove("legend")+rremove("ylab")+ theme(axis.text.x=element_text(size=8),axis.text.y=element_blank())

# define extra graphical objects
t1 <- textGrob("Lag 1")
t2 <- textGrob("Lead 1")
legend <- get_legend(lag_m)
blank <- grid.rect(gp=gpar(col="white"))

# plot interactions
png("figs/interactions.png", width = 8, height=4, units='in', res=300)
ggarrange(t1, lag_m+ rremove("legend"),lag_f1,lag_f2,blank,
          t2,lead_m+ rremove("legend"),lead_f1,lead_f2,legend, 
          ncol=5, nrow=2,widths = c(0.25,1,1,1, 0.5))
dev.off()
```




